### 2024-06-27 manuscript/manuscript.Rmd
Another edit pass

---
title: "A Comparison of Neural Network Models for Water Quality Forecasting"
authors:
  - name: Marcus Lapeyrolerie
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: mlapeyro@berkeley.edu
  - name: Carl Boettiger
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: cboettig@berkeley.edu (corresponding author)
abstract: |
  1. 
  2. 
  3. 
  4.

keywords:
  - Artificial Intelligence
  - Forecasting
  - Machine Learning
  - Time Series
  - Limnology

bibliography: references.bib
nocite: |

header-includes:
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{lineno}
  - \usepackage{bbm}
  - \usepackage{setspace}
  - \usepackage{graphicx}
  - \usepackage{subcaption}
  - \linenumbers
  - \doublespacing

output: 
  rticles::arxiv_article:
    keep_tex: true
  word_document: default
editor_options: 
  markdown: 
    wrap: sentence
---
\setlength{\parindent}{15pt}

# Introduction

The ability to accurately forecast water quality variables has become increasingly important in the era of global change. Freshwater ecosystems have been disproportionately impacted by anthropogenic activities, a trend that is expected to continue throughout the 21st century \cite{albert2021}. Since water quality variables, like water temperature, dissolved oxygen and chlorophyll-a concentration, influence many biological and physical processes in freshwater ecosystems, preemptive water quality forecasts could allow managers to evade situations with far-reaching negative consequences \cite{ouellet-proulx, stajkowski, chen_combining}. Limnologists have historically used statistical, process-based and machine learning models to forecast water quality metrics \cite{maier}; but, in recent years, researchers have shown that machine learning models are particularly well suited to take advantage of recent advancements in computer science and the rising availability of water quality data  \cite{hanson, zwart2023}. A shortcoming of past limnological forecasting studies that support the use of machine learning is that they tend to focus on a narrow selection of sites that are not representative of freshwater systems across a broad geographic scale. This paper has the primary aim of making a comprehensive comparison of state-of-the-art machine learning methods by evaluating forecasts at 34 different sites across North America at different times of the year.

The time series that are used in this study are taken from the National Ecological Observatory Network's (NEON) Ecological Forecasting Challenge, a competition where teams can submit forecasts for data that is collected and made publicly accessible by NEON \cite{thomas_span_2023}. A common finding across the challenge is that a day of year historical mean model (also referred to as the climatology model) commonly produces top scoring forecasts \cite{wheeler_predicting_2024, thomas_nearterm_2023}. For instance, in a model comparison that examined the forecasts for phenology, \cite{wheeler_predicting_2024} found that the climatology model outperformed all except one of the submitted models; and, the best performing model only marginally outperformed the climatology model. Thus, a primary consideration in this paper is comparing the performance of the machine learning models against the historical null model.

There are many promising neural network architectures that have not yet been evaluated for water quality forecasting. Most applications of neural networks to limnological time series have focused on the long-short-term-memory (LSTM) network, a model that was published in 1997 \cite{hochreiter_long_1997}. Over the nearly 30 years since LSTM's were introduced, researchers in computer science have built upon these earlier neural network architectures, introducing new models that achieve state of the art performance \cite{lim_temporal_2019, bai_empirical_2018, oreshkin_n-beats_2019}. Many recent studies that use neural networks for water quality forecasting have not explored more contemporary neural network architectures. We will address this gap in the literature by comparing 8 neural network models including LSTM's as well as more recently developed neural network models.

# Materials and Methods

We evaluated the forecasting performance of 12 different forecasting models on water temperature, dissolved oxygen and chlorophyll-a time series recorded by in-situ sensors at 34 freshwater sites across North America. These sites consist of Lakes, Non-wadeable Rivers and Wadeable Streams, subtypes classified by NEON. There is variability across sites in which target variables were observed. Additionally, maintenance issues led to gaps in the data which also varied across locations. Provided the lack of time series and large gaps at certain sites, we evaluated the forecast performance for water temperature, dissolved oxygen and chlorophyll-A at 33, 32 and 9 sites respectively.

The imputation of missing data proved to be critical to the performance of the ML models. After experimenting with data filling methods that resulted in poor model performance, we developed an imputation method inspired by the climatology model. If the gap size was less than 5 days, then the gap was filled using a Gaussian Process Filter. For gaps over 5 days, missing data was estimated using the daily historical mean when this statistic is available; and, when there are no data collected for a day of the year, either the monthly median, quarterly median, previous observation or global median was used in this order of preference. The intuition behind this method for missing data imputation is that this method will bias the neural networks towards the climatology model. Since it has been established that a climatology model produces top performing forecasts throughout the NEON forecasting challenge, backfilling with the climatology model seems likely to induce improved accuracy for the neural network models \cite{wheeler_predicting_2024, thomas_nearterm_2023}.

We compared the performance of the neural network models to the climatology and naive persistence model. The climatology model generates forecasts by finding the daily mean and standard deviation and draws samples from a Gaussian distribution with these parameters. The naive persistence model finds the last observed value from the target time series and predicts this value for each day in the forecast horizon. For each model, we computed the Continuous Ranked Probability Score (CRPS) and the Root Mean Square Error (RMSE). To gauge how models performed in relation to the climatology model, we computed the Continuous Ranked Probability Skill Score (CRPSS) which we defined as 
\begin{equation}
CRPSS_{model} = 1 - \frac{CRPS_{model}}{CRPS_{clim}}.
\end{equation}
Similarly, using the naive persistence model as a reference, we computed the RMSE Skill Score,
\begin{equation}
RMSE-SS_{model} = 1 - \frac{RMSE_{model}}{RMSE_{naive}}.
\end{equation}
If a target model outperforms the reference model, then the target model will have a positive skill score. If the target model performs worse than the reference model, the skill score will be negative.

## Method Group 1: Statistical Models (Theta)

To compare the performance of neural network models with a representative statistical model, we evaluated forecasts generated by the Theta model. Provided a seasonally adjusted univariate time series, the Theta model creates forecasts by modifying the second differences of the data \cite{assimakopoulos_theta_2000}. The magnitude of local curvature modifications is given by the Theta coefficient,

\begin{equation}
\nabla^2 Z_t(\theta) = \theta \, \nabla^2 X_t,
\end{equation}

where $\nabla$ is the difference operator, $Z_t$ is defined as a theta line and $X_t$ is the original data. When $\theta < 1$, the second differences are reduced from the data, yielding a Theta line that amplifies the long term trends of the data. For instance, if $\theta = 0$, the theta line will be a linear regression of the data. And if $\theta > 1$, the short term behavior of the data will be magnified in the Theta line. The Theta model generates a forecast by extrapolating the linear combination of two or more Theta lines. Although the Theta model is relatively simple, it has performed remarkably well in prominent forecasting competitions that include hard to predict data \cite{makridakis_m3-competition_2000, makridakis_m4_2020}. Since the Theta model was originally presented in 2000, new variations of the Theta model have been developed that outperform the original Theta model \cite{fiorucci_models_2016}. Throughout this study, we use the StatsForecast AutoTheta model which selects the best performing model from a range of Theta model variants.

## Method Group 2: Neural Network Models

All the machine learning algorithms that are used in this study are based on neural network architectures. Neural networks have the property of being universal function approximators, so it is possible, in theory, that all of these models could exactly approximate the data generating process \cite{hanin_universal_2019}. Yet, it is often the case that neural networks greatly underperform their function approximation capabilities in practice \cite{adcock_gap_2020}. This performance gap can be due to a variety of reasons including overfitting and insufficient hyperparameter tuning \cite{adcock_gap_2020}. The discrepancy between theory and practice in function approximation with neural networks motivates research on how machine learning models perform in specific domains.

The 8 machine learning models that we compare take a variety of approaches with the design of their neural networks. We will not go into detail on how these different models work, but for those interested to learn more, we list the models and their references in table \ref{}. An important concept to understand is that the neural network models considered in this study learn directly from the data and are not instilled with domain knowledge. This non-mechanisitic basis is at once very powerful as it does not restrict the models with misleading assumptions, but neural networks are also limiting in that they are not readily interpretable and often require more data than knowledge-guided or empirical methods to perform well \cite{karpatne_knowledge-guided_2024, read_processguided_2019}. For the management of critical resources like water, the lack of interpretability could be a deterrent to the adoption of NN methods, and there may not be enough data for some water systems to accommodate ML approaches \cite{zhi_deep_2024}.

The neural network models are configured in this study to provide estimates for predictive uncertainty. The neural network architectures that we use are deterministic, so they are not able to produce probabilistic forecasts intrinsically. We work around this limitation by performing quantile regression whereby the neural networks are trained to output quantiles at each time step in the forecast window. Probabilistic forecasts are then generated by drawing samples according to these quantiles.

For the 8 neural network models that we investigate, there are varying design choices made regarding the type of covariates that can be used. For this study, we employed 2 groups of models: one group used past covariates and the other used future covariates. For the models that accept past covariates (which included TCN, BlockRNN, NLinear, DLinear, Transformer and NBEATS), we used the other target variables recorded at that site as well as air temperature as covariates. For the models that only accept future covariates (which included TFT and RNN), we used the day of the year as the sole covariate. All the time series were split into training and validation sets, whereby the models were trained on time series from 2020 to 2023 and validated at 12 30-day non-overlapping intervals in 2023.

We provide fully reproducible code for fitting, scoring and visualizing the forecasts. All the machine learning forecasts were generated using the `darts` python library \cite{herzen_darts_2021}. `darts` is similar to other libraries like `scikit-learn` in Python or `tidymodels` in R in that the library allows users to employ a variety of time series forecasting models without having to implement them. The github repo for this study can be found at https://github.com/boettiger-lab/neon4cast-darts-ml.

# Results

We examine the forecast skill of 8 neural network models (RNN, BlockRNN, NBEATS, NLinear, DLinear, TFT, TCN and Transformer), 1 statistical model (AutoTheta) and 2 null models (naive persistence and climatology) on time series taken from the NEON Forecasting Challenge's Aquatics Theme. Following recent work that has established that multi-model ensembles offer advantages over individual models for water temperature forecasting \cite{olsson_multimodel_2024}, we were inspired to investigate multi-model ensembles in this comparative study. So, in addition to the ML, empirical and null models, we created an ensemble model that aggregates the forecasts taken from all the neural network models. For this neural network ensemble, all the NN models are represented equally, hence its name "Naive Ensemble".

In Table \ref{}, we present the mean CRPS and RMSE scores for the respective target variables. For dissolved oxygen (DO) and water temperature (WT), the models perform similarly: the neural network models outperform the AutoTheta, naive persistence and climatology model with few exceptions; and the naive ensemble model is the best performing model with respect to both CRPS and RMSE. Yet, with chlorophyll-a, there are different patterns in model performance: while the neural network models still generally outperform the reference models in CRPS, the Naive Ensemble model is no longer the top performing model in either CRPS or RMSE. Instead, BlockRNN and the naive persistence model attain the best performance with respect to CRPS and RMSE, respectively.

By examining some of the individual forecasts, as shown in Figure \ref{}, it is possible to gain intuition for why the neural network models perform well across the target variables. The AutoTheta model produces forecasts that resemble linear regressions based upon recently observed values. These relatively simple forecasts perform well for many evaluation intervals, but there a few cases when AutoTheta fails catastrophically, negatively impacting the model's overall performance. For instance, after a peak of dissolved oxygen in the winter of 2024, the AutoTheta model forecasts that DO will continue to increase which is opposed to the trend that DO peaks in the winter and declines through the spring. Similarly with Chlorophyll-A, the AutoTheta model wrongly extrapolates that a spike in chlorophyll-a will lead to higher chlorophyll-a concentrations instead of an immediate reversion to the non-bloom state.

Conversely, the neural network models are able to learn from historical trends. For instance, at the beginning of 2024, the neural network models have learned from the training data that DO peaks in the winter and declines through the spring. With chlorophyll-a, however, we see that the neural network models fail to predict any of the spikes in concentration which originate from algae blooms; instead, the neural networks take a conservative approach, only predicting that the chla concentration will remain at non-bloom levels throughout the year.

These general patterns in model performance can be observed in the skill score plots displayed in Figures \ref{}. The AutoTheta model tends to have longer tails, indicating that the AutoTheta forecasts are more likely to perform very well or very badly. Meanwhile, the neural network models generally have fewer forecasts that underperform relative to AutoTheta as well as fewer forecasts that are outlier outperformers. The distributions of NN skill scores are more skewed towards outperforming than AutoTheta's skill scores are. In these plots, we display scores according to water body type, but we did not see any significant differences in performance across these categories. However, this may be due to an underreporting of scores for lakes and non-wadeable rivers relative to the number of scores found for wadeable streams.

When examining the performance of the models within a forecast horizon, additional nuances emerge, confusing the perspective that neural network models are the best performing model class. In Figure \ref{}, the AutoTheta model is consistently the best performing model for short time horizons across target variables, but by the end of the 30-day horizon, the AutoTheta model is the worst perfoming model universally. Meanwhile the neural network models underperform AutoTheta during the early stages of the forecast window (generally, when t < 5), but their forecast skill declines less rapidly than the skill score of AutoTheta. So while, the neural networks generally outperform AutoTheta and the null models according to coarse scoring metrics like mean CRPS and RMSE, AutoTheta is the best performer in short horizons.

\begin{table}[!htbp]
    \centering
    \begin{subtable}{\textwidth}
        \centering
        \caption{Dissolved Oxygen ($\text{mg} \, \text{L}^{-1}$)}
        \begin{tabular}{lcc}
            \toprule
            Model & Mean CRPS & Mean RMSE \\
            \midrule
            Climatology & 0.62 & 0.92 \\
            Naive Persistence & & 1.55 \\
            AutoTheta & 0.67 & 1.02 \\
            BlockRNN & 0.52 & 0.80 \\
            DLinear & 0.52 & 0.81  \\
            NBEATS & 0.51 & 0.80 \\
            NLinear & 0.52 & 0.80 \\
            \textbf{NaiveEnsemble} & \textbf{0.47} & \textbf{0.74} \\
            RNN & 0.60 & 0.92 \\
            TCN & 0.62 & 0.98 \\
            TFT & 0.52 & 0.80 \\
            Transformer & 0.51 & 0.80 \\
            \bottomrule
        \end{tabular}
    \end{subtable}
    \begin{subtable}{\textwidth}
        \centering
        \caption{Water Temperature ($^{\circ} \text{C}$)}
        \begin{tabular}{lcc}
            \toprule
            Model & Mean CRPS & Mean RMSE \\
            \midrule
            Climatology & 1.46 & 2.19 \\
            Naive Persistence & & 6.05 \\
            AutoTheta & 1.55 & 2.35 \\
            BlockRNN & 1.45 & 2.26 \\
            DLinear & 1.35 & 2.08 \\
            NBEATS & 1.36 & 2.10 \\
            NLinear & 1.33 & 2.06 \\
            \textbf{NaiveEnsemble} & \textbf{1.18} & \textbf{1.80} \\
            RNN & 1.32 & 2.03 \\
            TCN & 1.89 & 3.02 \\
            TFT & 1.19 & 1.86 \\
            Transformer & 1.32 & 2.05 \\
            \bottomrule
        \end{tabular}
    \end{subtable}
    \begin{subtable}{\textwidth}
        \centering
        \caption{Chlorophyll-a ($\text{mg} \, \text{L}^{-1}$)}
        \begin{tabular}{lcc}
            \toprule
            Model & Mean CRPS & Mean RMSE \\
            \midrule
            Climatology & 4.83 & 7.50 \\
            \textbf{Naive Persistence} & & \textbf{4.86} \\
            AutoTheta &  4.20 & 6.31 \\
            \textbf{BlockRNN} & \textbf{3.14} & 5.16 \\
            DLinear & 3.47 & 5.48 \\
            NBEATS & 3.78 & 5.95 \\
            NLinear & 3.67 & 5.63 \\
            NaiveEnsemble & 3.44 & 5.50 \\
            RNN & 4.15 & 6.40 \\
            TCN & 3.72 & 5.77 \\
            TFT & 3.67 & 5.87 \\
            Transformer & 4.02 & 6.35 \\
            \bottomrule
        \end{tabular}
    \end{subtable}
    \caption{Mean CRPS and RMSE for dissolved oxygen, water temperature and chlorophyll-a forecasts. The neural network-based models generally outperform the statistical benchmark model, AutoTheta, as well as the null models, naive persistence and climatology, across the target variables. For dissolved oxygen and water temperature, the NN ensemble model is the best performing model with respect to CRPS and RMSE. With chlorophyll-a, BlockRNN attains the best CRPS score, and naive persistence attains the best RMSE.}
    \label{tab:mean_crps_rmse}
\end{table}

\begin{figure}[!htbp]
  \centering
  \begin{subfigure}{\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/sitetype_legend_0.png}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/type_local_historical_oxygen.png}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/type_local_naive_oxygen.png}
  \end{subfigure}
  \caption{Skill score distributions for dissolved oxygen. The CRPSS plot measures the CRPS skill relative to the climatology model. The RMSE-SS plot measures the RMSE skill relative to the naive persistence model. A guideline is plotted at the threshold for underperformance: a score above this line denotes that a forecast is underperforming the reference model; a score below this line denotes that a forecast is outperforming the reference model. The AutoTheta model exhibits longer tails relative to most of the neural network models, indicating that AutoTheta has a larger amount of forecasts that do either very well or very poorly. The neural network models tend to have shorter tails in the underperforming region.}
\end{figure}

\begin{figure}[!htbp]
  \centering
  \begin{subfigure}{\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/sitetype_legend_0.png}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/type_local_historical_temperature.png}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/type_local_naive_temperature.png}
  \end{subfigure}
  \caption{Skill score distributions for water temperature. AutoTheta does not have the long tails as seen in the skill score distributions for dissolved oxygen and chlorophyll-a, instead the neural network models have heavier tails in the outperforming region. This indicates that the neural network models are producing high accuracy forecasts more frequently than the AutoTheta model.}
\end{figure}

\begin{figure}[!htbp]
  \centering
  \begin{subfigure}{\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/sitetype_legend_1.png}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/type_local_historical_chla.png}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/type_local_naive_chla.png}
  \end{subfigure}
  \caption{Skill score distributions for chlorophyll-a. The neural network models have shorter tails in the underperforming region relative to AutoTheta, indicating that the neural network models are producing a smaller amount of inaccurate forecasts. The naive persistence null model attained the lowest RMSE on chlorophyll-a, and this is apparent in the RMSE-SS plot as the models have skill score distributions that are generally centered above the underperformance threshold.}
\end{figure}

\begin{figure}[!htbp]
  \centering
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/intra_historical_oxygen.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/intra_naive_oxygen.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \centering
    \rule{0pt}{0pt}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/intra_historical_temperature.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/intra_naive_temperature.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/intra_legend.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/intra_historical_chla.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/intra_naive_chla.png}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
    \centering
    \rule{0pt}{0pt}
  \end{subfigure}

  \caption{Mean skill scores within the 30-day forecast horizon. In the early phase of the forecast horizon (t < 5), the AutoTheta model is universally the top performing model. However, the performance of the AutoTheta model rapidly declines over the course of the 30-day horizon, concluding in AutoTheta being the worst performing model universally at the end of the horizon. The neural network models do not perform as well as AutoTheta at the beginning of the forecast horizon, but their performance declines less rapidly throughout the horizon.}
\end{figure}

\begin{figure}[!htbp]
  \centering
  \begin{subfigure}{\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/forecast_legend.png}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/oxygen_theta.png}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/chla_theta.png}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/oxygen_tft.png}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/chla_tft.png}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/oxygen_ensemble.png}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{./plots/chla_ensemble.png}
  \end{subfigure}
  \caption{Individual forecasts of dissolved oxygen and chlorophyll-a. The AutoTheta model produces accurate forecasts most of the year, especially for dissolved oxygen; but, there are some intervals where AutoTheta produces extremely inaccurate forecasts. After peaks in dissolved oxygen and chlorophyll-a, the AutoTheta model wrongly predicts that the values will continue to increase. Conversely, the neural network-based forecasts from the TFT and the Naive Ensemble model have learned from the training data that high levels of dissolved oxygen in the winter are followed by a steady decline throughout the spring. With chlorophyll-a, the neural networks take a conservative strategy where they predict that chlorophyll-a will remain in a non-bloom state, and these models never take a risk in forecasting that a bloom will occur.}
\end{figure}

# Discussion

Over the last decade, machine learning has gained significant prominence after a spate of high profile successes based on neural network models, some of these successes include beating the master champion of Go, accurately predicting protein structures and the advent of large language models like ChatGPT \cite{silver_mastering_2017, jumper_highly_2021, zhao_survey_2023}. During this time, researchers in computer science have developed new neural network models for time series forecasting with promises of significant accuracy improvements over previous methods \cite{makridakis_m4_2020}. Simultaneously during the last 10 years, there has been a growing focus in limnology towards using neural networks for water quality forecasting, but this focus has primarily centered on older neural network architectures witn little exploration of newer methods. This study aims to address this research gap by performing a comprehensive comparison of neural network models that include more recently developed models.

Instead of identifying that there is a new neural network model that performs exceptionally well, this study has affirmed the conventional wisdom that ensemble forecasts are often more accurate and robust than forecasts from individual models. We found that an ensemble model which aggregates the forecasts from all the neural network models was the best performing model overall, evincing all the other models on the target variables of dissolved oxygen and water temperature, while also performing well for chlorophyll-a, attaining the second and third best CRPS and RMSE respectively. This result affirms recent work from \cite{olsson_multimodel_2024} which established that multi-model ensembles offer advantages over individual models for water temperature forecasting.

Yet, as is often the case when judging whether one method is superior to another, we have found that subtle changes to our criteria may have produced radically different results. For instance, if we had use the same forecasts abbreviated to a 5 day horizon, then the AutoTheta model would have been the best performing model according to the scoring metrics used in this study. But, it is important to qualify that if we changed the model training configurations to consider shorter forecasts horizons, this would produce forecasts that will be different than the ones we evaluated for the 30-day horizon. Thus, how neural networks models perform over different forecast horizons warrants future research.

The neural network models struggled to forecast chlorophyll-a, a result that is not surprising given that forecasting algae blooms is well established as a challenging prediction problem in limnology \cite{chen_combining}. With cholorophyll-a, the neural network models displayed a tendency to take a conservative strategy, regularly predicting that the chla concentration would remain in a non-bloom state. It is possible that the neural network models settled on this conservative behavior because we did provide enough information to predict blooms; the neural networks were not provided with some of the typical covariates used in process-based models like Nitrogen and Phosphorous concentrations, and photosynthetic active radiation. Yet, it is also plausible that this conservative strategy was driven by the length of the forecast horizon. Blooms are stochastic events, characterized by rapid fluctuations in chlorophyll-a. Predicting the timing of such an event in a long term horizon will be more difficult than performing the same prediction on a shorter horizon as there are more opportunities in the long term horizon to be wrong. With this reasoning, it is sensible that the neural networks would take a conservative approach.

In place of the one question that we set out to answer, namely, "How do neural network models perform for water quality forecasting?", we have unearthed an abundance of more questions and ideas for future work. Outside of exploring how the performance of neural networks will vary with forecast horizon, some other promising directions for exploration include cross-learning and hybrid models. Throughout this study, we have trained individual models for each target time series. Cross-learning presents a different approach where we would train one model across all the time series available for a given target variable. Through cross-learning, the model might be able to learn a pattern from one time series that will transfer to making accurate forecasts on a different time series. We also reckon that hybrid models may have superior forecasting performance to neural network or statistical models individually. We observed that AutoTheta were the best performing models early in the forecast horizon, and neural networks performed best later in the forecast horizon. This motivates the exploration of a hybrid model that employs AutoTheta at the beginning of the forecast horizon and neural network models later in the forecast horizon.

It is an exciting time to study water quality forecasting. There are large, publicly accessible water quality data sets that exist now, and there will be an increasing amount of limnological data in the future as the cost for sensors and remote sensing continue to come down. Neural network models offer promising advantages over alternative methods for limnological forecasting as these methods can learn complex patterns from data without being restrained by a need for domain knowledge. Furthermore, their performance has the potential to improve as more water quality data becomes available. Yet, they also present a range of problems like a lack of interpretability and generalizability that could restrain them from being useful decision support tools for the safety critical problem of water resource management. This study does not attempt to provide a definitive answer as to whether neural networks will be the best method class moving forward as answering this question will take a plethora of research. Instead, we hope that this study will generate more ideas and incite momentum towards improving our ability to forecast and manage water quality.

# Acknowledgements

# Conflicts of Interest {-}

The authors declare there are no conflicts of interest.

# Authors' Contributions {-}

Marcus Lapeyrolerie and Carl Boettiger developed the code and wrote the manuscript.

# Data Availability {-}

\newpage

# References